Implementation of Full Synchronous Composition of Discrete Event Models Using IEC 61499 Function Blocks

### 2005-01-31 doc/Sync_Execution_with_IEC61499/paper.tex
*** empty log message ***

\author{\authorblockN{Goran Cengic, Knut Åkesson and
    Bengt Lennartson}
\authorblockA{
  Automation Laboratory\\
  Department of Signals and Systems\\
  Chalmers University of Technology\\
  Sweden
}
\and
\authorblockN{Chengyin Yuan and Placid Ferreira}
\authorblockA{
  Department of Mechanical and Industrial Engineering\\
  University of Illinois at Urbana-Champaign\\
  USA
}
}
\maketitle
\begin{abstract}
  A new method for implementation of synchronous composition
  is presented. Using an emerging standard, IEC 61499, the
  method enables efficient development of distributed
  control applications. It also brings together supervisory
  control theory with distributed control application
  development since synchronous composition of discrete
  event models is used to model interaction between systems
  in supervisory control theory. Method is demonstrated
  using a simple example before the formulation for the
  general case is presented. A new structure of control
  applications suitable for application of supervisory
  control theory to control application development is
  presented also.
\end{abstract}

@@

Shorter product life-cycles increase the demands on the
manufacturing system. Since the product, to be produced, in
the manufacturing plant is changing frequently the
production equipment needs to be modified too. The frequent
changes of the system make it time consuming and expensive
to update the control system application to reflect the
changes. The same problem applies to both system types
mentioned above and it becomes even larger in mission
critical control applications since the strict requirement
of guaranteeing the correct operation is imposed. Thus,
there is a need to introduce algorithms and tools to make it
simple and as safe as possible to modify the control system.
The supervisory control theory,\cite{rw:con:1989}, is a
promising framework that might be useful for this purpose.

@@

Synchronous composition, \cite{h:com:1985}, is a useful
approach to model the interaction between automata and is
used in this paper to model the interaction between the
system models.\\
\textbf{Definition}: Full Synchronous Composition of two DFAs\\
The full synchronous composition of two DFAs, $A^1$ and
$A^2$, is denoted ${A}^1||{A}^2=\langle
Q,\Sigma,\delta,q_i,Q_m\rangle$ where $Q=Q^1 \times Q^2$,
$\Sigma=\Sigma^1 \cup \Sigma^2$, $q_i=\langle {q^1_{i}},
{q^2_{i}} \rangle$, $Q_m={Q^1_{m}} \times {Q^2_{m}}$ and the
transition function is
\begin{displaymath}
  \begin{array}{l}
    \delta(\langle{q^1,q^2}\rangle,\sigma)=\\
    =\left\{ {\begin{array}{*{10}l}
          {{ \langle \delta ^1 (q^1 ,\sigma ), \delta ^2 (q^2 ,\sigma ) 
              \rangle}} &  \sigma \in \Gamma^1(q^1) \cap \Gamma^2(q^2) \\
          {{ \langle \delta ^1 (q^1 ,\sigma ), q^2 \rangle} } &  \sigma \in 
          \Gamma^1(q^1) \setminus \Sigma^2 \\
          { { \langle q^1, \delta ^2 (q^2 ,\sigma )\rangle}} & \sigma \in 
          \Gamma^2(q^2) \setminus \Sigma^1 \\
          { \rm undefined} & {{\rm otherwise.}}
        \end{array}} \right.
  \end{array}
\end{displaymath} 
where $q^1 \in Q^1$, $q^2 \in Q^2$ and $\sigma \in \Sigma$.
$\Box$ In other words, if an event is in both alphabets and
it is enabled in both automatons, the transitions associated
with it are made simultaneously in both automatons.
Otherwise, only the automaton containing the event in its
alphabet makes the transition. This definition can be
extended for several automata.\\
\textbf{Definition}: Full Synchronous Composition of $n$ DFAs\\
The full synchronous composition of DFAs, $A^1$, $A^2$ ...
$A^n$, is denoted ${A}^1||...||{A}^n=\langle
Q,\Sigma,\delta,q_i,Q_m\rangle$ where
\begin{eqnarray}
  Q & = & Q^1 \times Q^2 \times ... \times Q^n\nonumber\\
  \Sigma & = & \Sigma^1 \cup \Sigma^2 \cup ... \cup \Sigma^n\nonumber\\
  q_i & = & \langle q^1_i,q^2_i,...,q^n_i \rangle\nonumber\\
  Q_m & = & Q^1_m \times Q^2_m \times ... \times Q^n_m\nonumber
\end{eqnarray}
and the transition function is
\begin{multline}\label{eq:delta}
  \delta(\langle{q^1,...,q^n}\rangle,\sigma)=\\
  =
  \begin{cases}
    {\rm undefined} &{\rm if\ } \exists A^k:\sigma\in\Sigma^k\land\sigma\notin\Gamma^k(q^k)\\
    \langle q^{1+},\dots,q^{n+}\rangle & {\rm otherwise}
  \end{cases} 
\end{multline} 
where 
\begin{equation}\label{eq:qi}
  q^{k+}=
  \begin{cases}
    \delta^k(q^k,\sigma) & {\rm if\ }\sigma\in\Gamma^k(q^k)\\
    q^k                  & {\rm if\ }\sigma\notin\Sigma^k
  \end{cases}
\end{equation}
$\Box$\\
What this definition states is that in a group of automatons
that share one event if there is one automaton that does not
have that event enabled in its current state, than the event
is not allowed to occur in any of the other automatons in
the group. That is, one automaton is able to prevent all the
others from executing an event if it does not enable it at
the same time as all the other automatons. This property of
the definition is used to construct a supervisor which
prevents events from occurring. The concept of supervisor is
explained further in the next section which introduces the
supervisory control theory.

@@

Thus to control the system so that it satisfies the
specification, the synchronous composition of the plant and
supervisor models has to be implemented in the control
software. Using the conventional implementation methods,
i.e. implementing the synchronization using PLC programming
techniques, the solution becomes complex which makes it
difficult to distribute across multiple control computers.
Next section presents an example illustrating how the
synchronous composition can be implemented using the new
standard.

@@

Function block types that generate the behavior of the plant
and supervisor models for the example are shown in
Fig.~\ref{fig:systemFBs}. The event input of each block
type is \texttt{OCCURRED}. It signals that an discrete event
has occurred and that the model should update its state
according to the data inputs. The data inputs are boolean
variables representing the discrete events in the model's
alphabet. For well defined behavior only one of the data
inputs associated with the \texttt{OCCURRED} event must be
\texttt{true} when the event is received. The model's state
is then updated according to the discrete event that this
data input represents and the event output \texttt{DONE} is
generated. It signals that the model has updated its state
and that the data outputs are ready. The data outputs are
boolean variables representing the enabled events after the
model has updated its state, i.e. the events labeling the
transitions that are possible from the new state.

@@

\section{Synchronization: General Approach}
The method for implementation of the full synchronous
composition for the general system containing any number of
discrete event models is made up of two major steps. The
first step is creation of the function block types that
generate the behavior of the discrete event models. The
interface that these blocks have to adhere to is shown in
Fig.~\ref{fig:model_interface}. These blocks are
application specific but they can be automatically generated
from discrete event models.

>>>>

\author{\authorblockN{Goran Cengic, Knut Åkesson and
    Bengt Lennartson}
\authorblockA{
  Automation Laboratory\\
  Department of Signals and Systems\\
  Chalmers University of Technology\\
  Sweden
}
\and
\authorblockN{Chengyin Yuan and Placid Ferreira}
\authorblockA{
  Department of Mechanical and Industrial Engineering\\
  University of Illinois at Urbana-Champaign\\
  USA
}
}
\maketitle
\begin{abstract}
  A new method for implementation of synchronous composition
  is presented. Using an emerging standard, IEC 61499, the
  method enables efficient development of distributed
  control applications. It also brings together supervisory
  control theory with distributed control application
  development since synchronous composition of discrete
  event models is used to model interaction between systems
  in supervisory control theory. The method is demonstrated
  using a simple example before the formulation for the
  general case is presented. A new structure of control
  applications suitable for application of supervisory
  control theory to control application development is
  presented also.
\end{abstract}

@@

Shorter product life-cycles increase the demands on the
manufacturing system. Since the product, to be produced, in
the manufacturing plant is changing frequently the
production equipment needs to be modified too. The frequent
changes of the system make it time consuming and expensive
to update the control system application to reflect the
changes. The same problem applies to both system types
mentioned above and it becomes even larger in mission
critical control applications since the strict requirement
of guaranteeing, the correct operation is imposed. Thus,
there is a need to introduce algorithms and tools to make it
simple and as safe as possible to modify the control system.
The supervisory control theory,\cite{rw:con:1989}, is a
promising framework that might be useful for this purpose.

@@

Synchronous composition, \cite{h:com:1985}, is a useful
approach to model the interaction between automata and is
used in this paper to model the interaction between the
system models.\\
\textbf{Definition}: Full Synchronous Composition of two DFAs\\
The full synchronous composition of two DFAs, $A^1$ and
$A^2$, is denoted ${A}^1||{A}^2=\langle
Q,\Sigma,\delta,q_i,Q_m\rangle$ where $Q=Q^1 \times Q^2$,
$\Sigma=\Sigma^1 \cup \Sigma^2$, $q_i=\langle {q^1_{i}},
{q^2_{i}} \rangle$, $Q_m={Q^1_{m}} \times {Q^2_{m}}$ and the
transition function is
\begin{displaymath}
  \begin{array}{l}
    \delta(\langle{q^1,q^2}\rangle,\sigma)=\\
    =\left\{ {\begin{array}{*{10}l}
          {{ \langle \delta ^1 (q^1 ,\sigma ), \delta ^2 (q^2 ,\sigma ) 
              \rangle}} &  \sigma \in \Gamma^1(q^1) \cap \Gamma^2(q^2) \\
          {{ \langle \delta ^1 (q^1 ,\sigma ), q^2 \rangle} } &  \sigma \in 
          \Gamma^1(q^1) \setminus \Sigma^2 \\
          { { \langle q^1, \delta ^2 (q^2 ,\sigma )\rangle}} & \sigma \in 
          \Gamma^2(q^2) \setminus \Sigma^1 \\
          { \rm undefined} & {{\rm otherwise.}}
        \end{array}} \right.
  \end{array}
\end{displaymath} 
where $q^1 \in Q^1$, $q^2 \in Q^2$ and $\sigma \in \Sigma$.
$\Box$ In other words, if an event is in both alphabets and
it is enabled in both automatons, the transitions associated
with it are made simultaneously in both automatons.
Otherwise, only the automaton containing the event in its
alphabet makes the transition. This definition can be
extended for several automata.\\
\textbf{Definition}: Full Synchronous Composition of $n$ DFAs\\
The full synchronous composition of DFAs, $A^1$, $A^2$ ...
$A^n$, is denoted ${A}^1||...||{A}^n=\langle
Q,\Sigma,\delta,q_i,Q_m\rangle$ where
\begin{eqnarray}
  Q & = & Q^1 \times Q^2 \times ... \times Q^n\nonumber\\
  \Sigma & = & \Sigma^1 \cup \Sigma^2 \cup ... \cup \Sigma^n\nonumber\\
  q_i & = & \langle q^1_i,q^2_i,...,q^n_i \rangle\nonumber\\
  Q_m & = & Q^1_m \times Q^2_m \times ... \times Q^n_m\nonumber
\end{eqnarray}
and the transition function is
\begin{multline}\label{eq:delta}
  \delta(\langle{q^1,...,q^n}\rangle,\sigma)=\\
  =
  \begin{cases}
    {\rm undefined} &{\rm if\ } \exists A^k:\sigma\in\Sigma^k\land\sigma\notin\Gamma^k(q^k)\\
    \langle q^{1+},\dots,q^{n+}\rangle & {\rm otherwise}
  \end{cases} 
\end{multline} 
where 
\begin{equation}\label{eq:qi}
  q^{k+}=
  \begin{cases}
    \delta^k(q^k,\sigma) & {\rm if\ }\sigma\in\Gamma^k(q^k)\\
    q^k                  & {\rm if\ }\sigma\notin\Sigma^k
  \end{cases}
\end{equation}
$\Box$\\
What this definition states is that in a group of automatons
that share one event if there is one automaton that does not
have that event enabled in its current state, than the event
is not allowed to occur in any of the other automatons in
the group. That is, one automaton is able to prevent all the
others from executing an event if it does not enable it at
the same time as all the other automata. This property of
the definition is used to construct a supervisor which
prevents events from occurring. The concept of supervisor is
explained further in the next section which introduces the
supervisory control theory.

@@

Thus to control the system so that it satisfies the
specification, the synchronous composition of the plant and
supervisor models has to be implemented in the control
software. Using the conventional implementation methods,
i.e. implementing the synchronization using PLC programming
techniques, the solution becomes complex, which makes it
difficult to distribute across multiple control computers.
Next section presents an example illustrating how the
synchronous composition might be implemented using the new
standard.

@@

Function block types that generate the behavior of the plant
and supervisor models for the example are shown in
Fig.~\ref{fig:systemFBs}. The event input of each block
type is \texttt{OCCURRED}. It signals that an discrete event
has occurred and that the model should update its state
according to the data inputs. The data inputs are boolean
variables representing the discrete events in the model's
alphabet. For well defined behavior only one of the data
inputs associated with the \texttt{OCCURRED} event must be
\texttt{true} when the event is received. The model's state
is then updated according to the discrete event that this
data input represents and the event output \texttt{DONE} is
generated. It signals that the model has updated its state
and that the data outputs are ready. The data outputs are
boolean variables representing the enabled events after the
model has updated its state, i.e. the events that are possible from the new state.

@@

\section{Synchronization: General Approach}
The method for implementation of the full synchronous
composition for the general system, containing any number of
discrete event models, is made up of two major steps. The
first step is creation of the function block types that
generate the behavior of the discrete event models. The
interface that these blocks have to adhere to is shown in
Fig.~\ref{fig:model_interface}. These blocks are
application specific but they can be automatically generated
from discrete event models.
